---
title: "data extraction"
author: "Yousra"
date: "2025-04-25"
output: html_document
---



First create remote directory 
(base) yousra@UAVT16:~$ mkdir ~/remote_saillon
(base) yousra@UAVT16:~$ sshfs yousra@130.92.119.132:/data_1/saillon_24/metashape-projects/2024 ~/remote_saillon
The remote_saillon directory will be **created on local machine — it's just a folder that acts as a "window" into the remote server once it's mounted with sshfs. 


##function
Extract data Saillon 

```{r}
extract_reflectance_to_csv <- function(tif_folder, polygon_file, output_csv) {
  library(terra)
  library(sf)
  library(dplyr)
  library(stringr)

  # Load polygons
  polygons <- st_read(polygon_file)

  # List only reflectance orthomosaics
  tif_files <- list.files(tif_folder, pattern = "_Ortho\\.tif$", 
                          full.names = TRUE, recursive = TRUE)

  results <- list()

  for (tif_path in tif_files) {
    filename <- basename(tif_path)
    date_str <- str_extract(filename, "\\d{8}")
    date_fmt <- as.Date(date_str, format = "%Y%m%d")

    r <- rast(tif_path)

    # Skip files with unexpected number of bands
    if (nlyr(r) != 7) {
      message("Skipping file due to unexpected number of bands: ", filename)
      next
    }

    # Name and drop alpha band
    names(r) <- c("B", "G", "R", "RE", "NIR", "TIR", "alpha")
    r <- r[[names(r) != "alpha"]]

    # Match CRS
    if (!st_crs(polygons) == crs(r)) {
      polygons <- st_transform(polygons, crs(r))
    }

    # Define stat functions
    stat_functions <- list(mean = mean, median = median, sd = sd, min = min, max = max)

    # Extract stats
    extracted <- lapply(names(stat_functions), function(stat_name) {
      df <- terra::extract(r, vect(polygons), fun = stat_functions[[stat_name]], na.rm = TRUE) %>%
        as.data.frame()
      names(df)[-1] <- paste0(names(r), "_", stat_name)
      df
    })

    # Merge stats into one dataframe
    stats_df <- Reduce(function(x, y) merge(x, y, by = "ID"), extracted)

    # Clean up and annotate
    colnames(stats_df)[1] <- "polygon_id"
    stats_df$date <- date_fmt
    stats_df$image_name <- filename

    results[[length(results) + 1]] <- stats_df
  }

  # Combine all results
  reflectance_df <- bind_rows(results)

  # Save to CSV
  write.csv(reflectance_df, output_csv, row.names = FALSE)
  message("Reflectance data saved to: ", output_csv)

  return(reflectance_df)
}

```

extract data 
```{r}
tif_folder <- "/home/yousra/remote_saillon"
polygon_file <- "/home/yousra/Drone_drought_24/saillon_24/sp_saillon_1m.shp"
output_csv <- "/home/yousra/Drone_drought_24/reflectance_output.csv"

reflectance_data <- extract_reflectance_to_csv(tif_folder, polygon_file, output_csv)

```

rename polygons id to trees 

```{r}
# Rename 'polygon_id' to 'tree'
colnames(reflectance_data)[colnames(reflectance_data) == "polygon_id"] <- "Tree"

# Replace values in 'tree' column
reflectance_data$Tree <- factor(reflectance_data$Tree,
                                levels = c(1, 2, 3, 4),
                                labels = c("Beech1", "Beech2", "Oak1", "Oak2"))

# Output the modified data frame
print(reflectance_data)

```


##function
Extract data lens 

mkdir ~/remote_saillon
sshfs yousra@130.92.119.132:/data_1/lens_24/metashape-projects/2024 ~/remote_lens

```{r}
tif_folder <- "/home/yousra/remote_lens"
polygon_file <- "/home/yousra/Drone_drought_24/lens_24/lens_sp_1m.shp"
output_csv <- "/home/yousra/Drone_drought_24/reflectance_output_lens.csv"

reflectance_data_lens <- extract_reflectance_to_csv(tif_folder, polygon_file, output_csv)

```

rename polygons id to trees 

```{r}
# Rename 'polygon_id' to 'tree'
colnames(reflectance_data_lens)[colnames(reflectance_data_lens) == "polygon_id"] <- "Tree"

# Replace values in 'tree' column
reflectance_data_lens$Tree <- factor(reflectance_data_lens$Tree,
                                levels = c(1, 2),
                                labels = c("Pine1", "Pine2"))

# Output the modified data frame
print(reflectance_data_lens)

```

merge 2 data sets 
```{r}
reflectance_data_24 <- rbind(reflectance_data, reflectance_data_lens)
write.csv(reflectance_data_24, "/home/yousra/Drone_drought_24/reflectance_data_24.csv")

```

calculate vegetation indices 

```{r}

VI <- within(reflectance_data, {
  NDVI  <- (NIR_mean - R_mean) / (NIR_mean + R_mean)
  NDRE  <- (NIR_mean - RE_mean) / (NIR_mean + RE_mean)
  GNDVI <- (NIR_mean - G_mean) / (NIR_mean + G_mean)
  SR    <- NIR_mean / R_mean
  
  # Estimate CWSI from observed values (image-based method)
  T_min <- min(TIR_mean, na.rm = TRUE)
  T_max <- max(TIR_mean, na.rm = TRUE)
  CWSI  <- (TIR_mean - T_min) / (T_max - T_min)
})

```

Visualize data 

```{r}
library(ggplot2)

p1 <- ggplot(VI, aes(x = as.Date(date), y = NDVI, color = factor(tree))) +
  geom_line() +
  labs(title = "", x = "", y = "NDVI") +
  theme_bw()+
  theme(
    legend.title = element_blank())

p2 <- ggplot(VI, aes(x = as.Date(date), y = NDRE, color = factor(tree))) +
  geom_line() +
  labs(title = "", x = "", y = "NDRE") +
  theme_bw()+
   theme(
    legend.title = element_blank())

p3 <- ggplot(VI, aes(x = as.Date(date), y = GNDVI, color = factor(tree))) +
  geom_line() +
  labs(title = "", x = "", y = "GNDVI") +
  theme_bw()+
   theme(
    legend.title = element_blank())

p4 <- ggplot(VI, aes(x = as.Date(date), y = CWSI, color = factor(tree))) +
  geom_line() +
  labs(title = "", x = "", y = "CWSI") +
  theme_bw()+
   theme(
    legend.title = element_blank())

combined <- p1 + p2 + p3 + p4 + 
  plot_layout(ncol = 2) 
```

calculate vegetation indices 

```{r}

VI_24 <- within(reflectance_data_24, {
  NDVI  <- (NIR_mean - R_mean) / (NIR_mean + R_mean)
  NDRE  <- (NIR_mean - RE_mean) / (NIR_mean + RE_mean)
  GNDVI <- (NIR_mean - G_mean) / (NIR_mean + G_mean)
  SR    <- NIR_mean / R_mean
  
  # Estimate CWSI from observed values (image-based method)
  T_min <- min(TIR_mean, na.rm = TRUE)
  T_max <- max(TIR_mean, na.rm = TRUE)
  CWSI  <- (TIR_mean - T_min) / (T_max - T_min)
})

```

Visualize data 

```{r}
library(ggplot2)

p1 <- ggplot(VI_24, aes(x = as.Date(date), y = NDVI, color = factor(Tree))) +
  geom_line() +
  labs(title = "", x = "", y = "NDVI") +
  theme_bw()+
  theme(
    legend.title = element_blank())

p2 <- ggplot(VI_24, aes(x = as.Date(date), y = NDRE, color = factor(Tree))) +
  geom_line() +
  labs(title = "", x = "", y = "NDRE") +
  theme_bw()+
   theme(
    legend.title = element_blank())

p3 <- ggplot(VI_24, aes(x = as.Date(date), y = GNDVI, color = factor(Tree))) +
  geom_line() +
  labs(title = "", x = "", y = "GNDVI") +
  theme_bw()+
   theme(
    legend.title = element_blank())

p4 <- ggplot(VI_24, aes(x = as.Date(date), y = CWSI, color = factor(Tree))) +
  geom_line() +
  labs(title = "", x = "", y = "CWSI") +
  theme_bw()+
   theme(
    legend.title = element_blank())

combined <- p1 + p2 + p3 + p4 + 
  plot_layout(ncol = 2) 
```
